---
title: "postHocBinning"
author: "Alex M. Chubaty & Isolde Lane-Shaw"
date: "15 March 2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "hold")
```

# Overview

This module predicts expected bird densities based on land cover classification.

# Usage

```{r module_usage}
library(Require)
Require("SpaDES.core")

setPaths(modulePath = file.path(".."),
         inputPath = file.path("data"),
         outputPath = "outputs")

times <- list(start = 0, end = 10)

parameters <- list(
  postHocBinning = list()
)
modules <- list("postHocBinning")
objects <- list()
inputs <- list()
outputs <- list()

mySimOut <- simInitAndSpades(times = times, params = parameters, modules = modules, objects = objects)
```
# Parameters

Provide a summary of user-visible parameters.

```{r moduleParams, echo = FALSE}
df_params <- moduleParams("postHocBinning", "..")
knitr::kable(df_params)
```

# Events

Describe what happens for each event type.

## Plotting

Write what is plotted.

## Saving

Write what is saved.

# Data dependencies

## Input data

```{r moduleInputs, echo = FALSE}
df_inputs <- moduleInputs("postHocBinning", "..")
knitr::kable(df_inputs)
```

## Output data

Description of the module outputs.

```{r moduleOutputs, echo = FALSE}
df_outputs <- moduleOutputs("postHocBinning", "..")
knitr::kable(df_outputs)
```

# Links to other modules

Describe any anticipated linkages to other modules.

<!-- TODO: move everything below here into module code, e.g. defaults -->

```{r callLoad250mBootFunction}
birdsList <- c("BAWW", "OVEN") #specify the bird species
folderUrlBird <- "https://drive.google.com/drive/folders/1fCTr2P-3Bh-7Qh4W0SMJ_mT9rpsKvGEA" # give file location 

birdRasterStack <- load250mBootRasters(folderUrl = folderUrlBird,
                                       birdsList = birdsList,
                                       rastersPath = downloadFolderBird,
                                       rasterToMatch = landscapeRaster,
                                       studyArea = studyArea)

noBootsDFLocation <- "https://drive.google.com/file/d/1ldESo9gb6icRD8ZsuPgaEDSIwFjJEe4W"
noBootsDF <- drive_download(file = noBootsDFLocation, 
                            path = file.path(downloadFolderBird, "noBootsDF"),
                            type = "spreadsheet",
                            overwrite = TRUE)
```

# Create data table of bird density by landscape

A `data.table` is created, containing the number of cells of each land cover class that are found in the `landscapeRaster`, and the mean bird density for each land cover class, alongside the variance, and the standard error of this mean density. 

The first step is to gather the values from the rasters and create a clean dataset.
We then  calculate, for each cover class, the number of cells in this class, the mean bird density, and the variance and standard error for bird density. 

## Get bird dataset

Here the data for each cell's mean bird density and variance between bootstrap replicates is collected in a `data.table`.

```{r functionGetBirdDataset}
getBirdDataset <- function(birdRasterStack, categoriesRaster) {
  reproducible::Require("raster")
  
  meanBirdRasters <- names(birdRasterStack) %>%
    str_detect('mean') %>%
    keep(birdRasterStack, .)
  namesMeanBirdRasters <- names(meanBirdRasters)
  
  birdDatasets <- lapply(X = meanBirdRasters, FUN = function(birdRasterLayer) {
    landBirdRasterStack <- raster::stack(categoriesRaster, birdRasterLayer)
    ## take the values from the rasters and input them to a data table calledcellValues
    cellValues <- data.table(getValues(landBirdRasterStack))
    cellValues <- setnames(cellValues, c("landCoverClass", "birdDensity"))
    
    cellValues <- na.omit(cellValues) ## remove any rows with NA
    
    ## make landCoverClass categorical rather than numerical
    cellValues$landCoverClass <- as.factor(cellValues$landCoverClass) 
    
    return(cellValues)
  })
  
  names(birdDatasets) <- meanBirdRasters

  return(birdDatasets)
}
```

```{r getBirdDataset}
birdDatasets <- getBirdDataset(birdRasterStack = birdRasterStack,
                               categoriesRaster = landscapeRaster)

for (i in names(birdDatasets)) {
  attr(birdDatasets[[i]], "Species") <- i ## attr(birdDatasets$OVEN, "Species")
}
```

```{r getBirdStatsbyClass}
birdStatsByClass <- getBirdStatsByClass(birdDatasets = birdDatasets)

for (i in names(birdStatsByClass)) {
  attr(birdStatsByClass[[i]], "Species") <- i
}
```

```{r getAssumptionsSummary}
assumptionsSummary <- getAssumptionsSummary(birdStatsTables = birdStatsByClass)
```

```{r getKernelDensityData}
kernelDensityData <- getKernelDensityData(birdDatasets = birdDatasets)
```

```{r getKernelDensityPlot}
kernelDensityPlot <- getKernelDensityPlot(birdName = "Ovenbird", #sp name for title
                                          coverType = "7", #cover type name for title
                                          birdCoverDensity = kernelDensityData$OVEN$"7", #which kernel density cover type and bird species to plot
                                          meanData = birdStatsByClass$OVEN[7, meanBirdDensity] #show a red line with the mean density
)
#which sp and cover type to graph 
```

```{r plotVarBirdDensity}
# plotsVarBirdDensity <- lapply(X = birdStatsByClass,
#                               FUN = function(singleBirdStats){
#                                 
# plotVarBirdDensity <- ggplot(data = singleBirdStats, aes(x =landCoverClass, y = varBirdDensity)) + geom_bar(stat="identity")
# 
# return(plotVarBirdDensity)
# 
# })
```
